---
title: 'COVID-19: Analyze Mobility Trends'
output: html_document
---


```{r include=FALSE}
source("mobility.R", echo = FALSE, keep.source = TRUE)
```

The global lockdown has slowed down mobility considerably. This can be seen in the data produced by our ubiquitous mobile phones.
Apple is kind enough to make those anonymized and aggregated data available to the public. If you want to learn how to get a handle on those data and analyze trends with R read on!

To download the current data set go to the following website, click on “All Data CSV”: [Apple Maps: Mobility Trends Reports](https://www.apple.com/covid19/mobility) and move the file to your data folder.


To get an overview we first load the data into R and print the available regions (data for countries and many cities are available) and transportation types (“driving”, “transit” and “walking”):

```{r echo=TRUE}
mobility <- read.csv("data/applemobilitytrends-2020-04-21.csv") # change path and file name accordingly
levels(mobility$region)
```

```{r echo=TRUE}
levels(mobility$transportation_type)
```

We now create a function **mobi_trends** to return the data in a well-structured format. The default **plot = TRUE** plots the data, **plot = FALSE** returns a named vector with the raw data for further investigation:

```{r echo=TRUE}
mobi_trends <- function(reg = "United States", trans = "driving", plot = TRUE, addsmooth = TRUE) {
  data <- subset(mobility, region == reg & transportation_type == trans)[4:ncol(mobility)]
  dates <- as.Date(sapply(names(data), function(x) substr(x, start = 2, stop = 11)), "%Y.%m.%d")
  values <- as.numeric(data)
  series <- setNames(values, dates)
  if (plot) {
    plot(dates, values, main = paste("Mobility Trends", reg, trans), 
         xlab = "", ylab = "", type = "l", col = "blue", lwd = 3)
    if (addsmooth) {
      lines(dates, values, col = "lightblue", lwd = 3)
      lines(supsmu(dates, values), col = "blue", lwd = 2)
    }
    abline(h = 100)
    abline(h = c(0, 20, 40, 60, 80, 120, 140, 160, 180, 200), lty = 3)
    invisible(series)
  } else {
    series
  }
}

mobi_trends()
```

The drop is quite dramatic… by 60%! Even more dramatic, of course, is the situation in Italy:

```{r echo=TRUE}
mobi_trends(reg = "Italy")
```

A drop by 80%! The same plot for Frankfurt:

```{r echo=TRUE}
mobi_trends(reg = "Frankfurt")
```

Obviously in Germany people are taking those measures less seriously lately, there seems to be a clear upward trend. This can also be seen in the German “walking” data:

```{r echo=TRUE}
mobi_trends(reg = "Germany", trans = "walking")
```

What is interesting is that before the lockdown “transit” mobility seems to have accelerated before plunging:

```{r echo=TRUE}
mobi_trends(reg = "Germany", trans = "transit")
```

You can also plot the raw numbers only, without an added smoother (option **addsmooth = FALSE**):

```{r echo=TRUE}
mobi_trends(reg = "London", trans = "walking", addsmooth = FALSE)
```

And as I said, you can conduct your own analyses on the formatted vector of the time series (option **plot = FALSE**)…

```{r echo=TRUE}
mobi_trends(reg = "London", trans = "walking", plot = FALSE)
```

…as we have only scratched the surface of the many possibilities here, there are many interesting analyses, like including the data in epidemiological models or simply calculate correlations with new infections/deaths!

